\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-06 lang-design.tex
stuff

However, with \lsti{subst} being late bound, \lsti{subst} can no longer
be unfolded: the definition of \lsti{subst}, as a recursive function,
varies across families, yet the proof of the \lsti{ht_unit} case should be
reusable by a derived family regardless of how that family may add
new cases to the definition of \lsti{subst}.
How can the programmer change the goal to \lsti{hasty G tm_unit ty_unit}, then?

A key observation is that although late binding
prevents \emph{definitional equality} on \lsti{subst}, it does not
affect \emph{propositional equality}.
That is, \lsti{subst tm_unit x t' = tm_unit}, as a proposition
(\lsti{Prop}), can still be proved.
After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
from a base family to a derived family.
%
Hence, our plugin provides a tactic \lsti{fsimpl}
with which the programmer can rewrite \lsti{subst tm_unit x t'} to
\lsti{tm_unit}.

>>>>

However, with \lsti{subst} being late bound, \lsti{subst} can no longer
be unfolded (challenge~\ref{chg:latebind-v-defeq}):
the definition of \lsti{subst}, as a recursive function,
varies across families, yet a derived family should be able to reuse the
proof of the \lsti{ht_unit} case regardless of how it may add
new cases to the definition of \lsti{subst}.
Without the ability to unfold,
how can the programmer make progress in this proof, then?

A key observation is that although late binding
prevents \emph{definitional equality} on \lsti{subst}, it does not
affect \emph{propositional equality}.
That is, \lsti{subst tm_unit x t' = tm_unit}, as a proposition
(\lsti{Prop}), can still be proved.
After all, how \lsti{subst} is defined on \lsti{tm_unit} does not vary
from a base family to a derived family.
%
Based on this insight, our plugin automatically generates propositional
equalities for recursive functions defined using \lsti{FRecursion}.
It also provides a tactic \lsti{fsimpl} that enables, for instance,
simplifying \lsti{subst tm_unit x t'} to \lsti{tm_unit}.
Under the hood, \lsti{fsimpl} works by rewriting using the generated
propositional equalities.

Note that only within those families that contain "subst" is definitional
equality on \lsti{subst} unavailable.
Within those families, the meaning of \lsti{subst} is late bound; it is
polymorphic to the enclosing family.
In contrast, outside those families, \lsti{subst} can be referenced only
by specifying a family that contains it,
such as \lsti{STLC.subst} or \lsti{STLC_fix.subst}.
So Coq, by unfolding \lsti{STLC_fix.subst}, is able to consider
\lsti{STLC_fix.subst tm_unit x t'} and \lsti{tm_unit}
as definitionally equal.
